---
title: ""
author: "Jorge Luis Villalba"
date: "`r Sys.Date()`"
format: pdf
geometry:
  - left = 2.5cm
  #- textwidth=7in
  - top = 2.0cm
  - right = 2.5cm
header-includes:
  - \usepackage{fancyhdr} 
  - \usepackage{ragged2e} 
  - \usepackage{geometry} 
  - \usepackage{amsmath, amsfonts, amssymb, latexsym, amsthm, amsfonts} 
  - \usepackage{float} 
  - \usepackage{subcaption} 
  - \usepackage[dvips]{psfrag} 
  - \usepackage{enumerate} 
  - \usepackage{graphicx} 
  - \usepackage{caption} 
  - \usepackage{multirow}  % Para tablas
  - \usepackage{flushend} 
  - \usepackage{hyperref}

---

\renewcommand{\tablename}{Tabla}
\renewcommand{\figurename}{Figura}
\renewcommand{\refname}{References}

\renewcommand{\footrulewidth}{0.4pt}
\renewcommand{\headrulewidth}{0.4pt}
\hypersetup{
colorlinks=true,
linkcolor=blue,
filecolor=magenta,
urlcolor=cyan,
pdftitle={JORGE LUIS VILLALBA},
bookmarks=true,
pdfpagemode=FullScreen,
}


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, comment = FALSE, warning = FALSE)
```


\pagestyle{fancy}
\fancyhead[R]{}
\fancyhead[L]{\footnotesize{Caso de estudios }}
\fancyfoot[L]{\includegraphics[height=0.5cm]{imagenes/logoc.png}}
\fancyfoot[R]{}



\title{The style  
\includegraphics[width=1.3cm]{imagenes/logoR.png} + \LaTeX$^2$ ( Title in English)\\
El estilo \includegraphics[width=1.3cm]{imagenes/logoR.png} + \LaTeX$^2$  (Title in Spanish) }



\author{Jorge Villalba-Acevedo \thanks{ Ciencias básicas, Universidad Tecnológica de
Bolívar, Colombia, e-mail: \url{jvilllaba@utb.edu.co}}
\and Alejandro Gutiérrez-Vera \thanks{ Facultad de ingeniería, Universidad de Celaya, México, e-mail: \url{agutierrez@udec.edu.mx}}
\and  \thanks{ Ciencias básicas, Instituto Tecnológico de Celaya, méxico, \url{@udec.edu.mx}}
\and otro autor\thanks{}  
\and otro autor\thanks{}
}





\begin{@twocolumnfalse}
\maketitle

\begin{center}\rule{12cm}{0.3mm} \end{center}

\begin{center}
\textbf{Abstract} 
\end{center}
 
\justify

\textbf{Background:}

\textbf{Aims:} Jorge Luis Villalba 

\textbf{Methods:}

\textbf{Results:}

\textbf{Conclusions:}



\textbf{\textit{Keywords:}}
Colombia, méxico, word 3, word 4, word 5, word 6.


\begin{center}
\textbf{Resumen} 
\end{center}

\justify


\textbf{Antecedentes:}

\textbf{Objetivos:}

\textbf{Método:}

\textbf{Resultados:}
  
\textbf{Conclusión:}


\textbf{\textit{Palabras clave:}}
Colombia, méxico, word 3, word 4, word 5, word 6.


\begin{center}\rule{12cm}{0.3mm} \end{center}
\end{@twocolumnfalse}


## 1. Introduction

Distribution that, as in the preceding mode, can be distinguished by the occurrence probability allocated to the three scenarios.
 This paper presents an analysis of the distribution of the sample pro
 portion  and the diferene in sample proportions \cite{jorge}  taken from a randomly generated population textit Mi rosoft.
 

## 2. Data



Distribution that, as in the preceding mode, can be distinguished by the occurrence probability allocated to the three scenarios \cite{2}.
 This paper presents an analysis of the distribution of the sample pro
 portion  and the diferene in sample proportions  taken from a randomly generated population textit Mi rosoft \cite[p\'ag 21]{Baz}.
 
### Ejemplo de ecuación 1

\begin{equation}
\int x^3 dx = \dfrac{x^4}{4} + C
\label{intcub}
\end{equation} 



## 3. Materials and Methods



\begin{figure}
\centering
\includegraphics[height=3.5cm]{imagenes/logoc.png}
\caption{Villalba}
\end{figure}

 
## 4. Results and Discussion


Distribution that, as in the preceding mode, can be distinguished by the occurrence probability allocated to the three scenarios.


 This paper presents an analysis of the distribution of the sample pro
 
### Ejemplo de ecuación 2

$\textcolor{blue}{\alpha} = P(``\text{Error tipo I}")$

$\textcolor{blue}{\alpha}= P("\text{Rechazar} H_{0}" | "H_{0} \text{es verdadera}")$

$\textcolor{blue}{\alpha} = P(| \dfrac{\bar{x} - \theta_{0}}{\sigma/ \sqrt{n}}| > c  | \theta = \theta_{0} )$

$\textcolor{blue}{\alpha} = P(  |Z| > c )$ con $Z \sim\mathcal{N}(0,\,1)$.

### Ejemplo de gráfica 1


```{r, fig.height=3,fig.width=5,fig.cap="ggplot"}
library(tidyverse)
ggplot(iris, aes(x = Sepal.Width, y = Petal.Width, colour = Species))+
  geom_point()
```


\newpage

### Ejemplo de tabla 1



||**Pob1**| **Pob2**|**Pobext1**|**Pobext2**|**Gini1**|**Gini2**|
|---------------|------|------|------|------|------|------|
|               | 2010 | 2011 | 2010 | 2011 | 2010 | 2011 |
| Pasto         | 43,2 | 40,6 | 11,7 | 8,8  | 52,3 | 52,2 |
| Montería      | 39,7 | 37,5 | 6,7  | 6,5  | 52,5 | 53,0 |
| Barranquilla  | 39,5 | 34,7 | 7,4  | 5,3  | 49,7 | 47,2 |
| Cúcuta        | 39,3 | 33,9 | 8,4  | 5,7  | 47,9 | 47,1 |
| Cartagena     | 34,2 | 33,4 | 6,2  | 4,7  | 48,9 | 48,8 |
| Cali          | 26,1 | 25,1 | 6,4  | 5,2  | 52,9 | 50,4 |
| Villavicencio | 25,4 | 23,0 | 4,8  | 4,02 | 46,7 | 46,7 |
| Ibagué        | 26,6 | 22,0 | 4,3  | 2,7  | 49,5 | 44,9 |
| Pereira       | 26,8 | 21,6 | 3,8  | 2,2  | 45,6 | 45,1 |
| Manizales     | 23,8 | 19,2 | 4,7  | 2,3  | 49,5 | 47,1 |
| Medellín      | 22,0 | 19,2 | 5,6  | 4,0  | 53,8 | 50,7 |
| Bogotá        | 15,5 | 13,1 | 2,6  | 2,0  | 52,6 | 52,2 |
| Bucaramanga   | 10,9 | 10,7 | 1,2  | 1,1  | 45,0 | 44,9 | 

 : Datos de pobreza en Colombia en 2010 y 2011.


### Ejemplo de tabla 2

```{r}
knitr::kable(head(iris),caption = "marco de datos")
```



## 5. Conclusion and Recommendations

## Acknowledgments

## Appendix A. Dataset link

\newpage

\begin{thebibliography}{01}
\bibitem{jorge} Referencia n\'umero uno.
\bibitem{2} Referencia n\'umero dos.
\bibitem{3} Referencia n\'umero tres.
\bibitem{Baz} \textsc{Bazaraa, M.S., J.J. Jarvis} y \textsc{H.D. Sherali},
\textit{Programaci\'on lineal y flujo en redes}, segunda edici\'on,Limusa, M\'exico, DF, 2004.
\end{thebibliography}




