---
title: ""
author: ""
date: "`r Sys.Date()`"
format: pdf
geometry:
  - left = 2.5cm
  #- textwidth=7in
  - top = 2.0cm
  - right = 2.5cm
header-includes:
  - \usepackage{fancyhdr} 
  - \usepackage{ragged2e} 
  - \usepackage{geometry} 
  - \usepackage{amsmath, amsfonts, amssymb, latexsym, amsthm, amsfonts} 
  - \usepackage{float} 
  - \usepackage{subcaption} 
  - \usepackage[dvips]{psfrag} 
  - \usepackage{enumerate} 
  - \usepackage{graphicx} 
  - \usepackage{caption} 
  - \usepackage{multirow}  % Para tablas
  - \usepackage{flushend} 
  - \usepackage{hyperref}

---

\renewcommand{\tablename}{Tabla}
\renewcommand{\figurename}{Figura}
\renewcommand{\refname}{References}

\renewcommand{\footrulewidth}{0.4pt}
\renewcommand{\headrulewidth}{0.4pt}
\hypersetup{
colorlinks=true,
linkcolor=blue,
filecolor=magenta,
urlcolor=cyan,
pdftitle={Proyecto de Módulo},
bookmarks=true,
pdfpagemode=FullScreen,
}


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, comment = FALSE, warning = FALSE)
```


\pagestyle{fancy}
\fancyhead[R]{}
\fancyhead[L]{\footnotesize{Caso de estudios }}
\fancyfoot[L]{\includegraphics[height=0.5cm]{imagenes/logoc.png}}
\fancyfoot[R]{}




\title{
\LARGE
The Impact of SITM Vehicle Breakdowns on Economic Losses and Perceived Service Quality in Cartagena\\[0.5em]
\large % Cambia el tamaño aquí según lo que prefieras
Impacto de las varadas de vehículos del SITM en las pérdidas económicas y la percepción del servicio en Cartagena
}




\author{Alexander Sánchez González \thanks{ Universidad Tecnológica de
Bolívar, Colombia, e-mail: \url{sancheza@utb.edu.co}, ID: T00054514}
\and Amparo Hazbun Martinez \thanks{ Universidad Tecnológica de
Bolívar, Colombia, e-mail: \url{ahazbun@utb.edu.co}, ID: T000XXXXX}
\and Maria Mercedes Romero Racine \thanks{ Universidad Tecnológica de
Bolívar, Colombia, e-mail: \url{mracine@utb.edu.co}, ID: T000XXXXX}
\and Misael Jose Pastrana Fuentes \thanks{ Universidad Tecnológica de
Bolívar, Colombia, e-mail: \url{pastranam@utb.edu.co}, ID: T000XXXXX}
} 






\begin{@twocolumnfalse}
\maketitle

\begin{center}\rule{12cm}{0.3mm} \end{center}

\begin{center}
\textbf{Abstract} 
\end{center}
 
\justify

\textbf{Background:}

\textbf{Aims:} 

General Objective:

To analyze the impact of vehicle breakdowns within the Integrated System of Mass Transit (SITM) of the city of Cartagena on the economic losses of the operating company and on the perceived service quality by users.

Specific Objectives:

\begin{enumerate} 
  \item Identify the routes and locations with the highest frequency and costs of breakdowns. 
  \item Determine the main operational causes associated with breakdowns and their relationship with user satisfaction.    \item Quantify the economic losses caused by vehicle breakdowns in the SITM, taking into account operational and maintenance costs, as well as unproductive time. 
\end{enumerate}

\textbf{Methods:}

Analysis of breakdown records, analysis of economic losses, assessment of impact by route and location, statistical analysis.

\textbf{Results:}

\textbf{Conclusions:}



\textbf{\textit{Keywords:}}
Stranded, Economic losses, Service quality, Users, SITM (Integrated System of Mass Transit).


\begin{center}
\textbf{Resumen} 
\end{center}

\justify


\textbf{Antecedentes:}

\textbf{Objetivos:}

Objetivo general:

Analizar el impacto de las varadas de los vehículos del Sistema Integrado de Transporte Masivo (SITM) de la ciudad de Cartagena en las pérdidas económicas de la empresa operadora y en la calidad del servicio percibida por los usuarios.

Objetivos específicos:

\begin{enumerate}
    \item Identificar las rutas y ubicaciones con mayor frecuencia y costos de varadas.
    \item Determinar las principales causas operativas asociadas a las varadas y su relación con la satisfacción del usuario.
    \item Cuantificar las pérdidas económicas generadas por las varadas de los vehículos del SITM, considerando costos operativos, de mantenimiento y los tiempos improductivos.
\end{enumerate}

\textbf{Método:}

Análisis de registros de varadas, análisis de pérdidas económicas, evaluación del impacto por ruta y ubicación, análisis estadístico.

\textbf{Resultados:}
  
\textbf{Conclusión:}

\textbf{\textit{Palabras clave:}}
Varadas, Pérdidas económicas, Calidad del servicio, Usuarios, SITM (Sistema Integrado de Transporte Masivo).


\begin{center}\rule{12cm}{0.3mm} \end{center}
\end{@twocolumnfalse}


## 1. Introduction

The Integrated System of Mass Transit (SITM) in the city of Cartagena plays a fundamental role in urban mobility. Its proper functioning largely depends on the condition of the fleet and the operational efficiency of the vehicles. Consequently, one of the most recurring challenges faced by these systems is vehicle breakdowns or operational disruptions, which generate significant impacts for both the operating company and the users.

These breakdowns entail operational and maintenance costs associated with unproductive time. For users, such failures translate into delays, discomfort, and a loss of trust in the system.

This study aims to comprehensively analyze these failures within the SITM, starting with the identification of associated costs, as well as the routes and locations where they occur most frequently, in order to identify critical operational patterns.

Finally, the study seeks to quantify the economic losses resulting from breakdowns, highlight the main areas of impact, and assess the negative consequences for users, with the ultimate goal of enhancing transparency and rebuilding trust in the system.

Based on this context, the following research question arises:
What is the impact of vehicle breakdowns in Cartagena’s Integrated System of Mass Transit (SITM) on the operating company’s economic losses and on the perceived quality of service by users?

## 2. Data

The collected data corresponds to a representative sample of buses and routes within the city's Integrated Mass Transit System (SITM); therefore, it does not cover the system in its entirety. Additionally, the analysis is limited to a six-month period. The information was provided by one of the system’s operating companies. Both qualitative and quantitative data were gathered, although the latter predominates. Below is a detailed summary of the types and amounts of the final data to be used in the analysis.

```{r}
# Cargar las librerías necesarias
library(readxl)
library(dplyr)
library(knitr)
library(kableExtra)

# Cargar los datos desde el archivo Excel
data <- read_excel("../Dashboard_Presentacion/df_final.xlsx", sheet = "Sheet1")

# Obtener los tipos de variables
variable_types <- sapply(data, class)

# Calcular la cantidad de variables por tipo, ajustando las observaciones
num_quanti <- sum(sapply(data, is.numeric)) - 1  # Restamos year_falla, que ahora es Date/Time
num_quali <- sum(sapply(data, is.character)) - 2  # Restamos varado y afectacion_al_usuario, que son booleanas
num_dates <- sum(sapply(data, inherits, what = c("POSIXct", "POSIXt"))) + 1  # Agregamos year_falla
num_boolean <- 2  # Contamos varado y afectacion_al_usuario como booleanas

# Crear un resumen con la cantidad de variables por tipo
summary_table <- data.frame(
  Type = c("Quantitative", "Qualitative", "Date/Time", "Boolean"),
  Count = c(num_quanti, num_quali, num_dates, num_boolean)
)

# Mostrar el resumen en el PDF como tabla
kable(
  summary_table, 
  format = "latex", 
  caption = "Number of Variables by Type", 
  label = "summary-table"
) %>%
  kable_styling(position = "center", latex_options = "H")
```
```{r}
# Listar las variables por tipo
quanti_vars <- names(data)[sapply(data, is.numeric) & names(data) != "year_falla"]   # Variables cuantitativas
quali_vars <- names(data)[sapply(data, is.character) & !(names(data) %in% c("varado", "afectacion_al_usuario"))] # Variables cualitativas
date_vars <- c(names(data)[sapply(data, inherits, what = c("POSIXct", "POSIXt"))], "year_falla")  # Variables de tipo fecha
boolean_vars <- c("varado", "afectacion_al_usuario")  # Variables booleanas

# Crear una tabla con las variables por tipo y una breve descripción
variable_list <- data.frame(
  Type = c(
    rep("Quantitative", length(quanti_vars)),
    rep("Qualitative", length(quali_vars)),
    rep("Date/Time", length(date_vars)),
    rep("Boolean", length(boolean_vars))
  ),
  Variable = c(quanti_vars, quali_vars, date_vars, boolean_vars),
  Description = c(
    # Descripciones para variables cuantitativas
    "Cost associated with vehicle downtime",
    "Maximum number of trips made",
    "Minimum number of trips made",
    "Total number of users per route",
    "Delay time in minutes or hours",
    "Day when the vehicle was stranded",
    "Minimum number of passengers",
    "Maximum number of passengers",
    "Cost of downtime (minimum estimate)",
    "Cost of downtime (maximum estimate)",
    
    # Descripciones para variables cualitativas
    "Report System by the operating area",
    "Route number or identifier",
    "Location where the failure occurred",
    "Vehicle identifier or license plate",
    "Category or typology of the failure",
    "Mileage at the time of the failure",
    "Name of the vehicle operator",
    "Time of the incident",
    "Details or notes about the incident",
    "Decision taken regarding the failure",
    "Work order associated with the failure",
    "Responsible technician",
    "System involved in the failure",
    "Subsystem involved in the failure",
    "Reference component causing the failure",
    "Component affected",
    "Adjective describing the failure",
    "Consequence of the failure",
    "State or status of the failure resolution",
    "Additional observations",
    "Whether the day was a business day or not",
    "Month in which the failure occurred",
    "Time when it resumes operation.",
    
    # Descripciones para variables de tipo fecha/hora
    "Date when the failure occurred",
    "Datetime of the delay caused by the failure",
    "New date assigned after the failure",
    "Year in which the failure occurred",
    
    # Descripciones para variables booleanas
    "Whether the vehicle was stranded (Yes/No)",
    "Whether the user was affected (Yes/No)"
  )
)

# Mostrar la lista de variables en el PDF como tabla
kable(
  variable_list, 
  format = "latex", 
  caption = "List of Variables with Descriptions", 
  label = "variable-descriptions"
) %>%
  kable_styling(position = "center", latex_options = "H") 
```


## 3. Materials and Methods
 
## 4. Results and Discussion

## 5. Conclusion and Recommendations

## Acknowledgments

## Appendix A. Dataset link

\newpage

\begin{thebibliography}{01}
\bibitem{jorge} Referencia n\'umero uno.
\bibitem{2} Referencia n\'umero dos.
\bibitem{3} Referencia n\'umero tres.
\bibitem{Baz} \textsc{Bazaraa, M.S., J.J. Jarvis} y \textsc{H.D. Sherali},
\textit{Programaci\'on lineal y flujo en redes}, segunda edici\'on,Limusa, M\'exico, DF, 2004.
\end{thebibliography}




